Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Mechanical Perturbations

### 2024-08-02 main.tex
Remove todos, reduce table headers.

\subsection{Measurements}
%
We measure the time histories of the Bump'Em delivered perturbation forces and
the bicycle's steer angle, roll angle, roll angular rate, and forward speed.
Figure~\ref{fig:perturbation_0}\todo{Maybe remove the top plot of individual
forces and add plots of roll angle, roll rate, steer rate.} shows an example
perturbation force measurement compared to our Bump'Em controller command.
%
\begin{figure}
  \centering
  \includegraphics[width=140mm]{figures/torque_angle_perturbation_10.png}
  \caption{Motor force and resulting  steer torque perturbation based on an 80N
    counterclockwise applied force.}
  \label{fig:perturbation_0}
\end{figure}

@@

\section{Results}
%
The coefficient estimates for a single-level logistic regression at
6~\unit{\kph} with gain $g=10$ are shown in
Table~\ref{tab:freq-coefs-6}. The angular impulse, perturbation order, and
balance assist state are all statistically significant predictors. Larger
angular impulse increases the probability to fall and enduring more
perturbations or having the balance assist on, decrease the probability to
fall.  The associated multiplicative change in odds are shown in
Table~\ref{tab:freq-coefs-6}.
%
\begin{table}
  \centering
  \caption{Logistic regression coefficient estimates at
    6~\unit{\kilo\meter\per\hour} and gain $g=10$. Multiplicative change in
    odds of predictor variables at 6 \unit{\kilo\meter\per\hour} and gain
    $g=10$ based on frequentist single-level logistic regression.}
  \label{tab:freq-coefs-6}
  \begin{tabular}{lrrrrrr}
    \toprule
    Variable & Estimate & Standard error & \(p\)-value & MCiO & 2.5\% & 97.5\% \\
    \midrule
    Intercept & -0.29 & 0.17 & 0.09 & 0.75 & 0.53 & 1.05 \\
    Angular impulse & 1.69 & 0.27 & 0.00 & 5.40 & 3.18 & 9.16 \\
    Perturbation order & -0.77 & 0.22 & 0.00 & 0.46 & 0.30 & 0.72 \\
    Balance assist state & -0.64 & 0.27 & 0.02 & 0.53 & 0.31 & 0.89 \\
    Roll angle & -0.25 & 0.21 & 0.24 & 0.78 & 0.51 & 1.18 \\
    Steer angle & -0.14 & 0.21 & 0.51 & 0.87 & 0.58 & 1.32 \\
    Balance-assist on $\times$ roll angle & 0.52 & 0.34 & 0.12 & 1.68 & 0.86 & 3.29 \\
    Balance-assist on $\times$ steer angle & -0.41 & 0.34 & 0.22 & 0.66 & 0.34 & 1.28 \\
    Balance-assist on $\times$ angular impulse & 0.41 & 0.41 & 0.32 & 1.51 & 0.67 & 3.38 \\
    Balance-assist on $\times$ perturbation order & -0.53 & 0.34 & 0.12 & 0.59 & 0.30 & 1.15 \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{table}
  \centering
  \caption{Logistic regression coefficient estimates at
    10~\unit{\kilo\meter\per\hour} and gain $g=8$.  Multiplicative change in
    odds of predictor variables at 10 \unit{\kilo\meter\per\hour} based on
    frequentist single-level logistic regression. and 5\% confidence intervals
  of MCiO.}
  \label{tab:freq-coefs-10}
  \begin{tabular}{lrrrrrr}
    \toprule
    Variable & Estimate & Standard error & \(p\)-value & MCiO & 2.5\% & 97.5\% \\
    \midrule
    Intercept                                     & -0.24 & 0.16 & 0.13 & 0.78 & 0.57 & 1.07 \\
    Angular impulse                               & 2.39  & 0.29 & 0.00 & 10.92 & 6.23 & 19.13 \\
    Perturbation order                            & -1.16 & 0.21 & 0.00 & 0.31 & 0.21 & 0.48 \\
    Balance-assist on                             & -0.44 & 0.24 & 0.07 & 0.64 & 0.40 & 1.03 \\
    Roll angle                                    & 0.27  & 0.22 & 0.22 & 1.31 & 0.85 & 2.04 \\
    Steer angle                                   & -0.37 & 0.24 & 0.12 & 0.69 & 0.43 & 1.10 \\
    Balance-assist on $\times$ roll angle         & -0.61 & 0.34 & 0.08 & 0.54 & 0.28 & 1.07 \\
    Balance-assist on $\times$ steer angle        & 0.56  & 0.35 & 0.11 & 1.76 & 0.88 & 3.50 \\
    Balance-assist on $\times$ angular impulse    & 0.46  & 0.44 & 0.29 & 1.59 & 0.68 & 3.74 \\
    Balance-assist on $\times$ perturbation order & -0.37 & 0.32 & 0.25 & 0.69 & 0.37 & 1.30 \\
    \bottomrule
  \end{tabular}
\end{table}

\todo{Revisit if we want box plots to visualize the tables.}
Turning the balance-assist system on significantly \((p<0.05)\) reduces the
odds that a pertubation results in a fall occurs while cycling at a speed of
6~\si{\kph}. Figure~\ref{fig:probability-6kph} visualizes the probability of
falling as a function of the mean and centered angular impulse per participant
while keeping all other explantory factors at their centered mean values. This
figure is derived by setting all explanatory variables to there mean and
calculating the probality from Eq~\ref{eq:log-regress} for only change in
angular impulse given the estimates in Table~\ref{tab:freq-coeffs-6} which
indicate that the balance-assist system halves (0.53) the odds that a
perturbation results in a fall. This figure shows that for relatively large
impulses the probability to fall is unity for both states of balance assist on
and off. And for relatively small impulses the probability to fall is zero for
both states. But for impulses in the magnitude region (-1 to 0.5 STD) around
the mean-centered fall threshold, the probability of falling is significantly
lowered. The skewedness of the probaility curves arrives from the interaction
effects\todo[inline]{check this and think about it}.

>>>>

\subsection{Measurements}
%
We measure the time histories of the Bump'Em delivered perturbation forces and
the bicycle's steer angle, roll angle, roll angular rate, and forward speed.
Figure~\ref{fig:perturbation_0} shows an example perturbation force
measurement compared to our Bump'Em controller command.
%
\begin{figure}
  \centering
  \includegraphics[width=140mm]{figures/torque_angle_perturbation_10.png}
  \caption{Motor force and resulting  steer torque perturbation based on an 80N
    counterclockwise applied force.}
  \label{fig:perturbation_0}
\end{figure}

@@

\section{Results}
%
The coefficient estimates for a single-level logistic regression at
6~\unit{\kph} with gain $g=10$ are shown in
Table~\ref{tab:freq-coefs-6}. The angular impulse, perturbation order, and
balance assist state are all statistically significant predictors. Larger
angular impulse increases the probability to fall and enduring more
perturbations or having the balance assist on, decrease the probability to
fall.  The associated multiplicative change in odds are shown in
Table~\ref{tab:freq-coefs-6}.
%
\begin{table}
  \centering
  \caption{Logistic regression coefficient estimates at
    6~\unit{\kilo\meter\per\hour} and gain $g=10$. Multiplicative change in
    odds of predictor variables at 6 \unit{\kilo\meter\per\hour} and gain
    $g=10$ based on frequentist single-level logistic regression.}
  \label{tab:freq-coefs-6}
  \begin{tabular}{lrrrrrr}
    \toprule
    Variable & \(\alpha_k\) & SE & \(p\)-value & $e^{\alpha_k}$ & 2.5\% & 97.5\% \\
    \midrule
    Intercept & -0.29 & 0.17 & 0.09 & 0.75 & 0.53 & 1.05 \\
    Angular impulse & 1.69 & 0.27 & 0.00 & 5.40 & 3.18 & 9.16 \\
    Perturbation order & -0.77 & 0.22 & 0.00 & 0.46 & 0.30 & 0.72 \\
    Balance assist state & -0.64 & 0.27 & 0.02 & 0.53 & 0.31 & 0.89 \\
    Roll angle & -0.25 & 0.21 & 0.24 & 0.78 & 0.51 & 1.18 \\
    Steer angle & -0.14 & 0.21 & 0.51 & 0.87 & 0.58 & 1.32 \\
    Balance-assist on $\times$ roll angle & 0.52 & 0.34 & 0.12 & 1.68 & 0.86 & 3.29 \\
    Balance-assist on $\times$ steer angle & -0.41 & 0.34 & 0.22 & 0.66 & 0.34 & 1.28 \\
    Balance-assist on $\times$ angular impulse & 0.41 & 0.41 & 0.32 & 1.51 & 0.67 & 3.38 \\
    Balance-assist on $\times$ perturbation order & -0.53 & 0.34 & 0.12 & 0.59 & 0.30 & 1.15 \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{table}
  \centering
  \caption{Logistic regression coefficient estimates at
    10~\unit{\kilo\meter\per\hour} and gain $g=8$.  Multiplicative change in
    odds of predictor variables at 10 \unit{\kilo\meter\per\hour} based on
    frequentist single-level logistic regression. and 5\% confidence intervals
  of MCiO.}
  \label{tab:freq-coefs-10}
  \begin{tabular}{lrrrrrr}
    \toprule
    Variable & \(\alpha_k\) & SE & \(p\)-value & $e^{\alpha_k}$ & 2.5\% & 97.5\% \\
    \midrule
    Intercept                                     & -0.24 & 0.16 & 0.13 & 0.78 & 0.57 & 1.07 \\
    Angular impulse                               & 2.39  & 0.29 & 0.00 & 10.92 & 6.23 & 19.13 \\
    Perturbation order                            & -1.16 & 0.21 & 0.00 & 0.31 & 0.21 & 0.48 \\
    Balance-assist on                             & -0.44 & 0.24 & 0.07 & 0.64 & 0.40 & 1.03 \\
    Roll angle                                    & 0.27  & 0.22 & 0.22 & 1.31 & 0.85 & 2.04 \\
    Steer angle                                   & -0.37 & 0.24 & 0.12 & 0.69 & 0.43 & 1.10 \\
    Balance-assist on $\times$ roll angle         & -0.61 & 0.34 & 0.08 & 0.54 & 0.28 & 1.07 \\
    Balance-assist on $\times$ steer angle        & 0.56  & 0.35 & 0.11 & 1.76 & 0.88 & 3.50 \\
    Balance-assist on $\times$ angular impulse    & 0.46  & 0.44 & 0.29 & 1.59 & 0.68 & 3.74 \\
    Balance-assist on $\times$ perturbation order & -0.37 & 0.32 & 0.25 & 0.69 & 0.37 & 1.30 \\
    \bottomrule
  \end{tabular}
\end{table}

Turning the balance-assist system on significantly \((p<0.05)\) reduces the
odds that a pertubation results in a fall occurs while cycling at a speed of
6~\si{\kph}. Figure~\ref{fig:probability-6kph} visualizes the probability of
falling as a function of the mean and centered angular impulse per participant
while keeping all other explantory factors at their centered mean values. This
figure is derived by setting all explanatory variables to there mean and
calculating the probality from Eq~\ref{eq:log-regress} for only change in
angular impulse given the estimates in Table~\ref{tab:freq-coeffs-6} which
indicate that the balance-assist system halves (0.53) the odds that a
perturbation results in a fall. This figure shows that for relatively large
impulses the probability to fall is unity for both states of balance assist on
and off. And for relatively small impulses the probability to fall is zero for
both states. But for impulses in the magnitude region (-1 to 0.5 STD) around
the mean-centered fall threshold, the probability of falling is significantly
lowered. The skewedness of the probaility curves arrives from the interaction
effects\todo[inline]{check this and think about it}.
